% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file. 

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.

% Updated July 2018: Text block size changed from 6.5" to 7"

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{int,graphicx,endnotes,url}
\setcounter{secnumdepth}{4}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Moonshot: Implementing the Next Generation of Network
Telemetry Technologies}

%for single author (just remove % characters)
\author{
{\rm Andy Gospodarek}\\
Broadcom Corporation\\
gospo@broadcom.com
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle

\subsection*{Abstract}
Current network monitoring and telemetry applications require host-based
collectors across all nodes in a network (servers/hypervisors
and traditional switches/routers). These can be effective solutions,
but just as datacenter deployment patterns have evolved new technology
to track traffic as it moves through the network has emerged. Newer
specifications like Inband Network Telemetry\cite{INT} (INT) and Inband
Flow Analyzer\cite{IFA} (IFA) propose standards to add metadata to
packets or clone and add metadata as they flow through a network to
allow collectors/agents to gather data at the network edges. Hardware
that supports INT/IFA can add metadata automatically with
application/flowlevel/virtual-port granularity which allows more
detailed network monitoring and assurance to customers that service
levels for applications are being met.

\section{Introduction to Network Telemetry}

INT and IFA are all designed create a generic method of reporting and
collecting network state information on individual flows as the packets
traverse a network.  This allows for collection of data from individual
hosts or applications as frames that are part of those flows are marked
with \textit{telemetry headers} as they entry a \textit{telemetry
domain}.  Network devices can interpret telemetry header fields as
\textit{telemetry instructions} and a capable device will update packet
headers and header-fields \textit{in-situ} -- as a frame traverses the
network.  Marking frames as they travel through a network allows
detailed reporting of the exact path used by packets on the
network as well as enables real-time feedback loops and event detection.
This information can also be sent to an external collector for
post-processing if desired.  

\subsection{Network Telemetry Components}

Despite using slightly different nomenclature, the fundamental
components of the telemetry technolgies covered in the paper
are similar.

\subsubsection{Source or Initiator Node}
This is a trusted entity that creates the initial telemetry header and
instructions and places them into packets that are transmitted.  One
important note is that this node is expected to only add an initial
telemetry header to a small, sampled percentage of the traffic
transmitted.

\subsubsection{Transit Hop or Transit Node}
Any network element that adds telemetry metadata to a packet that
that contains supported telemetry header and instructions.

\subsubsection{Sink or Terminating Node}
This is a trusted entity that removes all telemetry headers,
instructions, and metadata from in-situ frames or drops out-of-band
frames to make the existence of telemetry applications transparent to
applications.  This trusted entity will use the headers and other local
configuration to determine if information needs to be sent to a
collector.  This node will then format and deliver the frame to the
Collector node or application.

\subsubsection{Collector}
An application that will receive telemetry data collected by a Sink or
Terminating node.

\subsubsection{Telemetry Domain}
A set of interconnected devices participating in telemetry actions all
under the same administrative control.  It is critical that devices in
the same domain are all configured to behave in a consistent manner.

\subsubsection{Typical Packet Flow}

Below is a diagram of a typical packet path of a network flow and IFA
flow through the components described in the previous section.  In this
case the IFA flow is a sample of the flow, so two frames travel between
the Initiator Node and the Terminating Node.

This was adapted from the latest IFA specification at the time of this
writing:
\tiny
\begin{center}
\begin{verbatim}
                              +----------+
                              |          |
                              |Collector |<-------------+
                              |          |              |
                              +----------+              |
                                                        |
                                                        |
                                                        |
                                                        |
                                                        |
                                                        |
                                                        | 
                                                        |
     +--------------+        +------------+        +----+-----------+
     |Initiator Node|        |Transit Node|        |Terminating Node|
flow |   +------+   |  flow  |  +------+  |  flow  |     +------+   | flow
---->|   | IFA  |   |------->|  | IFA  |  |------->|     | IFA  |   |---->
     |   +------+   |IFA flow|  +------+  |IFA flow|     +------+   |
     +--------------+(cloned)+------------+(cloned)+----------------+

     |------------------------ IFA Domain --------------------------|

\end{verbatim}
\end{center}
\normalsize

Initiator and Terminating Nodes could be servers with supported hardware
and software stacks or switches, routers, or other forwarding elements
on a network.  The flows may originate or teminate on the IFA/INT node
if that node is on the edge of the IFA Domain.  If the network extends
beyond the IFA Domain, flows originate or terminate outside the IFA
Domain.

This paper does not intend to cover the full scope of each telemetry
technology and feature; anyone who would like to learn more should
consider reading the latest INT and IFA specifications.  This paper will
cover some basic frame formats of each proposal in order to provide
context for an implementation discussion.

\section{Inband Network Telemetry (INT)}

Inband Network Telemetry is a framework suggested by
those interested in using P4 to create a programmable pipeline for
networking forwarding elements.  INT has multiple methods for collecting
information about the network:

\begin{itemize}
\item A sampled percentage of the frames are updated as they traverse
the network 
\item New \textit{probe packets} are injected and used to collect
telemetry information as they traverse the network.
\end{itemize}
In addition to defining the frame format and fields, the latest INT
specification also conveniently provides a P4 program specification for
INT Transmit for hardware or software that can use a P4 datapath.
Support for INT is not limited to devices using a programmable datapath
like P4.

\subsubsection{INT Frame Format}

The current INT specification describes the following formats as being able to
support additional encapsulation headers to support INT:

\begin{itemize}
\item INT over VXLAN (as VXLAN payload, per GPE extension)
\item INT over Geneve (as Geneve option)
\item INT over GRE (as a shim between GRE header and encapsulated payload)
\item INT over NSH (as NSH payload)
\end{itemize}

Additionally the INT specification also describes how DSCP bits or
\textit{probe markers} can be placed in the payload of packet (after the Layer4
header) to support these packet formats.

\begin{itemize}
\item INT over TCP (as payload)
\item INT over UDP (as payload)
\end{itemize}

Though many datacenter networks use encapsulated traffic (VXLAN, Geneve,
or GRE), the fact that INT does not have native support unencapsulated
traffic (standard IPv4/IPv6 and TCP/UDP) could be an issue for some
deployments.

The specified frame format for an INT IPv4/TCP frame would be as follows:

\begin{center}
\tiny
\begin{verbatim}
       0                   1                   2                   3
       0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                        IP Header                              |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                         Layer 4                               |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                      INT Shim Header                          |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                    INT Metadata Header                        |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                         Payload                               |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\normalsize
\end{center}

Remember that the INT Headers and Payload together are viewed as the full
payload to any non-INT-aware device, so anytime INT headers are added to
a packet any fields that account for the size of the packet or
payload will need to be adjusted.

If the decision to use a reserved DSCP mark (0x17 in this case) to
indicate a packet contained INT headers would cause the IPv4 header to
look like this:

\tiny
\begin{center}
\begin{verbatim}
       0                   1                   2                   3
       0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
      IPv4 Header:
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      | Ver=4 | IHL=5 | DSCP=0x17 |ECN|          Total Length         |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |         Identification        |Flags|      Fragment Offset    |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |  Time to Live | Protocol = TCP|         Header Checksum       |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                     Source IPv4 Address                       |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                  Destination IPv4 Address                     |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{center}
\normalsize

The INT specification also outlines suggestions for how to deal with
frames as they grow beyond the MTU, how to deal with false detection of
\textit{probe markers} contained in payload data of non-INT frames, as well as
other deployment scenarios.

\section{Inband Flow Analyzer (IFA)}

The initial IFA specification was drafted later than other initial
telemetry technologies and while similar, it aims to address some of the
shortcomings of INT.  One of the main differences is the
ability to send telemetry metadata out-of-band via cloned frames rather
than via the original datagram.  Like INT however, IFA supports adding
metadata to live traffic in-band.  

Allowing cloned frames provides benefits over in-band modification of
frames.  One benefit of cloning is administrators do not need to be
concerned about frames growing beyond the MTU.  As expected frame
payloads will be truncated when needed.

The proposed frame/header format was modified significantly from the INT
specification.  The goal was to make the frame format more acceptable to
devices that were not IFA-aware.

\subsubsection{IFA Frame Format}
The IFA spec outlines a significantly different scheme for the location
to telemetry metadata.  From the start IFA aims to interoperate with
encapsulated and unencapsulated IPv4 and IPv6 traffic.  This is accomplished by
using the IPv4 \textit{Protocol} and IPv6 \textit{Next Header} fields to
specify that this frame is an IFA frame.  There is no current IANA
reservation for IFA protocol, so testing should use one of the
experimental protocol numbers as described by RFC3692\cite{RFC3692}.

One specified frame format for an IFA IPv4/TCP frame would be as follows:
\tiny
\begin{center}
\begin{verbatim}
       0                   1                   2                   3
       0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                        IP Header                              |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                       IFA Header                              |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                         Layer 4                               |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                   IFA Metadata Header                         |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                   IFA Metadata Stack                          |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                         Payload                               |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{center}
\normalsize

A closer look at the IPv4 header demonstrates that Protocol=IFA would be
used to signal that this frame is an IFA frame:

\tiny
\begin{center}
\begin{verbatim}
       0                   1                   2                   3
       0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
      IPv4 Header:
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |Version|  IHL  |Type of Service|          Total Length         |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |         Identification        |Flags|      Fragment Offset    |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |  Time to Live | Protocol = IFA|         Header Checksum       |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                     Source IPv4 Address                       |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |                  Destination IPv4 Address                     |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{center}
\normalsize

Additionally the IFA Header provides a Next Header field that would
indicate that TCP is the next protocol:

\tiny
\begin{center}
\begin{verbatim}
       0                   1                   2                   3
       0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
      IFA Header:
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
      |Ver=2.0|  GNS  | NextHdr = TCP |R|R|R|M|T|I|T|C|   Max Length  |
      |       |       |               | | | |F|S| |A| |               |
      +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{center}
\normalsize

\section{Linux Kernel Software Architecture}

In order to add support for a software model for INT or IFA, it is
important to understand the discrete componets available within the
Linux ecosystem that could be used with or without modification to
add support for INT or IFA.  What follows is a list of each of the
nodes in a telemetry implementation, the functional blocks needed to
implement those features, and what existing kernel components could be
the target for performing those functions.

\subsection{Initiator Node}
The critical datapath operations of sample and mirror (IFA) or sample
and redirect(INT/IFA) are the most important part of the Initiator Node
implementation.

\begin{table}[h!]
  \begin{center}
    \caption{Initiator Node}
    \label{tab:table1}
    \begin{tabular}{l|l}
      \textbf{Desired Functionality} & \textbf{Kernel Implementation} \\
      \hline
      Packet sampling & TC or netfilter \\
      \hline
      Packet mirror(IFA) & TC \\
      \hline
      Packet redirect(INT/IFA) & TC or XDP/eBPF \\
      \hline
      Encapsulate and TX INT/IFA & lwtunnel or XDP/eBPF \\
    \end{tabular}
  \end{center}
\end{table}


\subsection{Transit Node}
For a device to function as a Transit Node there must be support to
recognize the frame in the kernel datapath.  Once the frame is
recognized one could either pass the traffic to a lwtunnel or eBPF
program for modification.

\begin{table}[h!]
  \begin{center}
    \caption{Transit Node}
    \label{tab:table2}
    \begin{tabular}{l|l}
      \textbf{Desired Functionality} & \textbf{Kernel Implementation} \\
      \hline
      Match on INT/IFA Packet & TC \\
      \hline
      Update metadata and TX & lwtunnel or XDP/eBPF \\
    \end{tabular}
  \end{center}
\end{table}

\subsection{Terminating Node}
The ability to match on all of the formats (or the enabled formats used
in a deployment) is the critical role of a Terminating Node.  Without this
functionality in TC or netfilter a system running the Linux kernel would
not be able to serve as a collector.
\begin{table}[h!]
  \begin{center}
    \caption{Terminating Node}
    \label{tab:table3}
    \begin{tabular}{l|l}
      \textbf{Desired Functionality} & \textbf{Kernel Implementation} \\
      \hline
      Match on INT/IFA frame & TC or netfilter \\
      \hline
      Extract metadata from frame & lwtunnel or XDP/eBPF \\
      \hline
      Send frame to collector & lwtunnel or XDP/eBPF \\
      \hline
      Transmit original frame (INT) & lwtunnel or XDP/eBPF \\
    \end{tabular}
  \end{center}
\end{table}


\subsection{Implementation Summary}
The largest volume of work to support INT/IFA in the Linux kernel
is included in the TC (specifically related to sampling) and lwtunnel
subsystem.  While an initial proof-of-concept could be done with eBPF,
long-term it seems like use of the lwtunnel infrastructure would be
better.

Use of the kernel for performing INT/IFA encapsulation could be seen by
some as more work than necessary.  Some would argue that this could be
solved with a simple userspace daemon.  The need to support
in-band encapsulation of packets means that the Linux kernel
must be aware of this encapsulation.

The desire to add support to configure hardware that is capable of
INT/IFA offload also means that this functionality \textbf{must} be
included in the Linux kernel.

\section{Beyond the Linux Kernel Datapath}

\subsection{Hardware Support for Telemetry}

As with any new standards-based technology, manufacturers of networking
hardware will make hardware capable of INT and/or IFA when there is
significant customer demand.  Existing devices with programmable
pipelines seem like the primary target, but fixed function devices could
also support for IFA/INT if there is enough demand.

\subsection{Network-based configuration}

There are also use-cases where datacenter administrators will want to
enable IFA on a server host without requiring communication with the
server OS.  This could be done via configuration commands on a BMC or
management processor or directly with a NIC.  While this method is
traditionally seen as less popular to those within the open-source
community who expect to have exclusive control over their hardware,
industry deployment trends indicate that this is a popular and prefereed
method for many.

The main use-case for network-based configuration is in datacenters
where baremetal servers are provided to users.  In these deployments
network adminstrators still desire the ability to interact more directly
with network devices.  Servers acting as an Source or Initiator Node
could make the lives of network adminstrators easier.  As NIC hardware
that supports INT/IFA emerges configuration in this manner will gain
popularity.

A RESTful API could be used to control the INT/IFA configuration
of a server much the same way that such APIs are often used to control
traditional networking equipment.  There may ultimately be a common
framework that exists to enable INT/IFA on a server or switch with the
same API.

\subsection{Complete Userspace Implementations}

No discussion about new datapath technologies would be complete without
addressing whether or not this could be implemented with a poll mode
driver like DPDK\cite{DPDK}.  It is unlikely that anyone would implement
or deploy a DPDK-based application on a server or switch for the sole
purpose of performing any of the functions needed for INT/IFA.

It is likely, however, that someone using a DPDK-based application for
performing forwarding in the core of a network may want to add INT/IFA
functionality to their application.  A library that could create, add
metadata, or terminate telemetry packets could be written to integrate
with any dataplane applcation.  At the time of this writing no known
library or solution exists.

\section{Telemetry Development and Deployment}

\subsection{User Risk}

Obviously the deployment of any new technology brings along with it some
risk.  Even looking at the changes between the first two versions of
IFA\footnote{https://tools.ietf.org/rfcdiff?url2=draft-kumar-ippm-ifa-01.txt} 
highlights that this is an evolving standard with the addition of new
metadata formats and an increased list of packet formats.  One way to
minimize risk is to be sure that any devices participating are all
managed/controlled by the same organization.

\subsection{Community Risk}

One of the major risks to adding new code to the Linux kernel or any
dataplane application is what happens to existing code when standards
change.  For users of upstream kernels or who build their own
applications this is not that difficult -- a \texttt{git pull} and
\texttt{make ...} can result in a code updated to the latest standard
supported by that project.

Unfortunately a standard like INT or IFA will may also require changes
to other network infrastructure to handle these changes and not all
administrators may have access to the latest hardware running the latest
versions of software.  As mentioned in the previous section, unfied
management of infrastructure will prevent his from being a major issue.

\section{Acknowledgments}

Special thanks to all those invoved in IFA and INT specs.  Thanks also
to those at Broadcom who did not particiapte in the writing of the
standard, but did provide code, feedback, etc during this process.

{\normalsize \bibliographystyle{acm}
\bibliography{int}}

\end{document}
